---
title: "Activity 9: Dates and times"
output: pdf_document
date: "2024-02-27"
---

```{r, echo = F}

# load packages
library(tidyverse)
library(lubridate)
library(ggridges)
```

## Getting data from the internet, using lubridate, learning a new plot.

The goal of this exercise is to use our skills to get data from online, transform the date/time, and learn how to make a new plot (ridgeline plot). The data we will use for this exercise is the biketown data.

### Question 1:

* Get the data about rides in January, July, and November of 2017, and **bind** them into a single data frame using `bind_rows`:

(Hint: use the `readr::read_csv` function to directly read the csv from the url pointing to the data)

```{r}

janBike <- "https://s3.amazonaws.com/biketown-tripdata-public/2017_01.csv"
julyBike <- "https://s3.amazonaws.com/biketown-tripdata-public/2017_07.csv"
Novbike <- "https://s3.amazonaws.com/biketown-tripdata-public/2017_11.csv"

JanData <- readr::read_csv(janBike) 

JulData <- readr::read_csv(julyBike) 

NovData <- readr::read_csv(Novbike) 

biketown_data <- bind_rows(JanData,JulData, NovData )



biketown_data <- biketown_data %>%
  select(StartDate, StartTime, EndDate, EndTime, Distance_Miles,
         BikeID)

glimpse(biketown_data)


```


### Question 2:

* Reproduce the plot provided in the PDF version of this activity, using the `lubridate` functions(`mdy`, `ymd_hms`, `wday`, `month`):

```{r}
mutate(StartDate = mdy(StartDate),
         EndDate = mdy(EndDate)) %>%
  mutate(StartDateTime = ymd_hms(paste(StartDate, StartTime, sep = " ")),
         EndDateTime = ymd_hms(paste(EndDate, EndTime, sep = " "))) %>%
  filter(Distance_Miles < 10) # filter for outlying trips!
```

```{r, eval = F, fig.width = 12, fig.height = 8}

biketown_dt %>%
  mutate(day_of_week = wday(StartDateTime, label = TRUE),
         month = month(StartDateTime, label = TRUE)) %>%
  group_by(day_of_week, month) %>%
  ggplot(aes(y = day_of_week, x = Distance_Miles, fill = factor(day_of_week))) +
  geom_density_ridges(scale = 4) +
  facet_wrap(~ month, nrow = 3) +
  labs(x = 'Distance (Miles)', y = 'Day of the week') +
  theme_ridges() +
  theme(legend.position = 'none')

```
